
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Virtual Analog Modeling}
\label{chapter:virtual_analog_modeling}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Acf{VA} models are digital emulations of audio systems that were originally built using analog electronic or electromechanical components \cite{Zoelzer2011}. They arose out of the demand for the reproduction of characteristic tonal distortions of these devices but with the digital stability and the ease of integration with existing software. Devices modelled range from filters through time-varying effects, amplifiers, and mechanical reverb units to tape or vinyl distortions.

One can distinguish three general approaches to \ac{VA} modeling \cite{Kiiski2016,Wright2020}. In \emph{black-box} modelling, only the input-output relation of a system is examined and a signal model is constructed to mimic that behavior. Neural networks have successfully been applied to this kind of modeling for guitar amplifiers \cite{Wright2019,Wrightetal2020}. 

In \emph{white-box} modelling the internal structure of the system under study is examined and used to construct an algorithm aiming at reproducing its behavior. Sometimes it is referred to as \emph{physical modeling}. Typical approaches in this category are a numerical solution of \acp{ODE} derived from electronic circuit analysis \cite{Yeh2007,Eichas2014} or wave-digital filters \cite{PASPWEB2010}. 

\emph{Grey-box} modeling falls somewhere in between the two already mentioned approaches, where we use some knowledge about the inner workings of the device under study to design a model and, subsequently, take advantage of the input and output data to adjust the model's parameters. This approach has been successfully applied to time-varying effects modeling like a phaser or a flanger \cite{Kiiski2016,Wright2020}, where an \ac{LFO} signal is estimated and used for conditioning the model.

The aim of this work was to investigate the applicability of the ODENet architecture to \ac{VA} modeling of audio effects. The main goal was to model a phaser pedal. However, in order to get intuition behind this type of architecture, a simpler, time-invariant system was modeled, namely, the first-order diode clipper. The following section describes this system and current research into its modeling. The subsequent section deals with the phaser in a similar manner. Finally, \Section{subsec:va_evaluation} describes evaluation methods of \ac{VA} models.

\input{tex/chapter_va/diode_clipper.tex}
\input{tex/chapter_va/phaser.tex}
\input{tex/chapter_va/evaluation.tex}
